A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market

### 2013-08-30 reports/Draft.tex
continued progress on draft

\usepackage{alltt}
\usepackage[round]{natbib}
\usepackage{caption, amsmath, graphicx, hyperref}

@@

In this section we look for evidence of temporal non-stationarity in the hedonic coefficient on traffic noise within our data. Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable ``\texttt{post}'' for all sales after September 2008. We then regress our LWR noise coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}, 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*month} + \alpha _2  \texttt{*post} + \alpha _3 \texttt{*post*month} + \epsilon.
\end{equation}
The results of this regression will help us understand the patterns of change in the implicit price of traffic noise. In particular, $\hat{\alpha}_1 \neq 0$ suggests a linear temporal trend in the marginal willingness to pay to avoid traffic noise, while $\hat{\alpha}_2 \neq 0$ suggests that there was a structural break after September 2008, and $\hat{\alpha}_3 \neq 0$ will imply a change in the temporal trend after September 2008. The linear regression results presented in Table \ref{tab:betaMAX} show that there is a significant negative shock to the estimated LWR coefficients for traffic noise after September 2008. Additionally, the coefficients also begin to a significant negative trend.
\begin{table}[h]
\caption{Regression Results: Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{verbatim}
            Estimate Std. Error t value Pr(>|t|)   
(Intercept) -1.7e-03   4.1e-05   -43.4   < 2e-16 ***
Month        6.9e-06   2.0e-06     3.5  0.000498 ***
Post        -8.0e-04   7.2e-05   -11.1   < 2e-16 ***
Month*Post  -8.9e-05   4.5e-06   -19.7   < 2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

@@

\begin{table}[ht]
\begin{center}
\caption{A Comparison of the Noise Regression Coefficient Estimates in Dakota County with and without Additional Structural Variables}\label{tab:Dak}
\begin{tabular}{lllc}
 & \multicolumn{2}{p{1.25in}}{Additional Variables Included in LWR?} & \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} & \\ \cline{1-3}
Model (1) = Structural Variables & -0.0015 & -0.0015 & mean\\ 
   & (0.0023) & (0.0031) & (st.\ dev.) \\ 
Model (2) = Model (1) + Locational Variables & -0.0014 & -0.0014  & \\ 
   & (0.0018) & (0.0022) &  \\ 
Model (3) = Model (2) + City Fixed Effects & -0.0014 & -0.0014 & \\ 
   & (0.0022) & (0.0018) & \\ 
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:Dak} shows that the traffic noise coefficients estimated by our LWR models are similar regardless of whether the additional structural variables were included or not. Welch two-sample t-tests fail to reject the null hypothesis of zero mean differences across our three models (comparing LWR estimates with vs. without the additional structural variables included in the model). Paired t-tests find differences for Model 2 and 3, but the estimated differences are zero to four decimal places and in one case the noise impact estimates with the additional variables are slightly larger and in the other case they are slightly smaller. We also conducted simple linear regressions of the noise coefficients without the additional structural variables on the noise coefficients obtained with the additional structural variables for each of our three models. In all cases the intercept estimates were close zero with slopes almost exactly equal to one and each also had $R^2$ values over 0.7, signifying that the hedonic coefficients for the traffic noise variable are unaffected by omitted variable bias from common structural variables. While we cannot confirm that this lack of ommitted variable bias also occurs elsewhere in our data, the results from Dakota County are promising.

@@

Contrary to the previous results presented by \citet{MarmolejoDuarteCarlos;GonzalezTamez2009} and \citet{Theebe2004a}, we find little evidence to suggest that the impact of traffic noise varies over space or by level of noise within our study area. The traffic noise coefficient was one of only two variables with regression coefficient standard deviations smaller than or similar to the simulated distributions obtained from Monte Carlo experiments. We do, however, find significant temporal variation in the impact of traffic noise in our data. The estimated impact of one additional decibel of traffic noise is a 0.19 percent reduction in the sale price of houses before September 2008, vs. 0.37 percent after September 2008. Lastly, after controlling for the Great Recession, we find no significant differences in the impact of traffic noise across the range of commonly seen noise levels (50-70 dB).

>>>>

\usepackage{alltt}
\usepackage[round]{natbib}
\usepackage{caption, amsmath, graphicx, hyperref, setspace, multirow, color}

@@

In this section we look for evidence of temporal non-stationarity in the hedonic coefficient on traffic noise within our data. Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable ``\texttt{post}'' for all sales after September 2008. We then regress our LWR noise coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}, 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*Month} + \alpha _2  \texttt{*Post} + \alpha _3 \texttt{*Post*Month} + \epsilon.
\end{equation}
The results of this regression will help us understand the patterns of change in the implicit price of traffic noise. In particular, $\alpha_1 \neq 0$ suggests a linear temporal trend in the marginal willingness to pay to avoid traffic noise, while $\alpha_2 \neq 0$ suggests that there was a structural break after September 2008, and $\alpha_3 \neq 0$ will imply a change in the temporal trend after September 2008. The linear regression results presented in Table \ref{tab:betaMAX} show that there is a significant negative shock to the estimated LWR coefficients for traffic noise after September 2008. Additionally, the coefficients also begin to a significant negative trend.
\begin{table}[h]
\caption{Regression Results: Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{verbatim}
            Estimate Std. Error t value Pr(>|t|)   
(Intercept) -1.7e-03   4.1e-05   -43.4   < 2e-16 ***
Month        6.9e-06   2.0e-06     3.5  0.000498 ***
Post        -8.0e-04   7.2e-05   -11.1   < 2e-16 ***
Month*Post  -8.9e-05   4.5e-06   -19.7   < 2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

@@

\begin{table}[ht]
\begin{center}
\caption{A Comparison of the Noise Regression Coefficient Estimates in Dakota County with and without Additional Structural Variables}\label{tab:Dak}
\begin{tabular}{lllc}
 & \multicolumn{2}{p{1.25in}}{Additional Variables Included in LWR?} & \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} & \\ \cline{1-3}
\multirow{2}{*}{Model (1) = Structural Variables} & -0.0015 & -0.0015 & mean\\ 
   & (0.0023) & (0.0031) & (st.\ dev.) \\[.2cm]
\multirow{2}{*}{Model (2) = Model (1) + Locational Variables} & -0.0014 & -0.0014  & \\ 
   & (0.0018) & (0.0022) &  \\[.2cm] 
\multirow{2}{*}{Model (3) = Model (2) + City Fixed Effects} & -0.0014 & -0.0014 & \\ 
   & (0.0022) & (0.0018) & \\ 
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:Dak} shows that the traffic noise coefficients estimated by our LWR models are similar regardless of whether the additional structural variables were included or not. Welch two-sample t-tests fail to reject the null hypothesis of zero mean differences across our three models (comparing LWR estimates with vs.\ without the additional structural variables included in the model). Paired t-tests find differences for Model 2 and 3, but the estimated differences are zero to four decimal places and in one case the noise impact estimates with the additional variables are slightly larger and in the other case they are slightly smaller. We also conducted simple linear regressions of the noise coefficients without the additional structural variables on the noise coefficients obtained with the additional structural variables for each of our three models. In all cases the intercept estimates were close zero with slopes almost exactly equal to one and each also had $R^2$ values over 0.7, signifying that the hedonic coefficients for the traffic noise variable are unaffected by omitted variable bias from common structural variables. While we cannot confirm that this lack of ommitted variable bias also occurs elsewhere in our data, the results from Dakota County are promising.

@@

Contrary to the previous results presented by \citet{MarmolejoDuarteCarlos;GonzalezTamez2009} and \citet{Theebe2004a}, we find little evidence to suggest that the impact of traffic noise varies over space or by level of noise within our study area. The traffic noise coefficient was one of only two variables with regression coefficient standard deviations smaller than or similar to the simulated distributions obtained from Monte Carlo experiments. We do, however, find significant temporal variation in the impact of traffic noise in our data. The estimated impact of one additional decibel of traffic noise is a 0.19 percent reduction in the sale price of houses before September 2008, vs.\ 0.37 percent after September 2008. Lastly, after controlling for the Great Recession, we find no significant differences in the impact of traffic noise across the range of commonly seen noise levels (50-70 dB).
